Learning Reusable Components for Neurally--Guided Bayesian Program Learning

### 2018-05-12 nips_2018.tex
shortened and got rid of a lot of wake/sleep

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Start from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
because it iterates between three different phases: an
\textbf{Explore} phase uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} phase
modifies the DSL by discovering regularities in the programs found by
the previous Explore phase; and a \textbf{Compile} phase, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
network ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).

 \begin{wrapfig}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfig}

@@

\\\noindent \textbf{Wake: Searching for programs.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we find more programs solving the tasks.
\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce the DSL
from the programs found in the wake phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on richer DSLs that
more closely match the domain.\footnote{This is loosely biologically
    inspired by the formation of abstractions during sleep memory
    consolidation~\cite{DUDAI201520}}
\\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during waking. As the DSL
improves and as search finds more programs, the recognition model gets
both more data to train on, and better data.\footnote{These two sources are also
    loosely biologically inspired by the distinct episodic replay and
    hallucination components of dream sleep~\cite{fosse2003dreaming}}

%% Our goal  is to induce a DSL while finding programs solving each of the tasks.
%% Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
%% to accelerate the search over programs -- and (3) improving the DSL.
%% The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
%% \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
%% When these improve, we find more programs solving the tasks.
%% \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure. As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.

%% The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
%% learning a generative model and training a recognition model on samples from the generative model.
%% We apply this idea to program induction, motivating why we call our algorithm the \system.

\begin{wrapfig}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.

\subsection{Probabilistic Framing}\label{mathematicalFraming}
Viewed as a probabilistic inference problem,
our goal is to jointly infer a single DSL (written $\mathcal{D}$)
and many programs (written $p$), one for each set of inputs/outputs (\emph{tasks}, written $x$).
A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a distribution over programs.
Fig.~\ref{graphicalModel}(a) diagrams this inference problem as a hierarchical Bayesian model.
Inference in this model is
difficult because the programs are unobserved,
and so we must solve a hard search problem to recover them. To make
search tractable we learn a bottom-up \emph{recognition
  model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
The recognition model $q(\cdot )$ is a neural
network that regresses from input/output pairs to a distribution over
programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
inference
scheme~\cite{le2016inference}.
 The neural recognition model and the
generative model embodied in the DSL jointly train each other, as they
iteratively learn to solve more programming tasks.

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL $\mathcal{D}$.

We frame this problem as maximum a posteriori (MAP) inference in the
generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model
during the Sleep-R phase:

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
Rather than directly predicting a distribution over $p$ conditioned on $x$,
the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
This taps into the intuition that programming is primarily a top-down activity:
as human programmers, we often first decide what kind of
programming constructs we might need to use,
and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

\subsection{Wake Phase: Searching for Programs}\label{explorationSection}

@@

\subsection{Sleep-R: Learning a Recognition Model}\label{recognitionSection}

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are probable
under $(\mathcal{D},\theta)$ while also assigning high likelihood for a task
according to $\probability[x|p]$.

The recognition model $q$ is a neural network that predicts,
for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
Together with the DSL, this defines a distribution over programs,
$\probability[p|\mathcal{D},\theta = q(x)]$.
We abbreviate this distribution as $q(p|x)$.
The crucial aspect of this framing is that the neural network
leverages the structure of the learned DSL,
so it is \emph{not} responsible for
generating programs wholesale.
We share this aspect with DeepCoder~\cite{balog2016deepcoder}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
\begin{align}
\mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
\mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
\mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}

@@

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
for which we have effective approaches like the Inside/Outside algorithm~(Lafferty,~2000), %\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.

@@

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We perform a wake cycle before each of the sleep cycles.
(2) On the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a generative model has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
%% (3) Because the frontiers can grow very large,
%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
(4) For added robustness, we enumerate programs from both the generative model and the recognition model.

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     %     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Wake})}
     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Sleep-R}})
     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Wake)}}

     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Sleep-G)}}
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D},\theta,q$
\end{algorithmic}
\end{algorithm}

>>>>

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Starting from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
because it iterates between three different phases: an
\textbf{Explore} phase uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} phase
modifies the DSL by discovering regularities in the programs found by
the previous Explore phase; and a \textbf{Compile} phase, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfigure}

@@

\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we find more programs solving the tasks.
\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
from the programs found in the exploration phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on richer DSLs that
more closely match the domain.
\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during exploration. As the DSL
improves and as search finds more programs, the recognition model gets
both more data to train on, and better data.
\begin{wrapfigure}{r}{0.6\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.

\subsection{Probabilistic Framing}\label{mathematicalFraming}

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a generative model over programs.
We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta],\quad
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta, \quad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}

%% Inference in this model is
%% difficult because the programs are unobserved,
%% and so we must solve a hard search problem to recover them. To make
%% search tractable we learn a bottom-up \emph{recognition
%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
%% The recognition model $q(\cdot )$ is a neural
%% network that regresses from input/output pairs to a distribution over
%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
%% inference
%% scheme~\cite{le2016inference}.
%%  The neural recognition model and the
%% generative model embodied in the DSL jointly train each other, as they
%% iteratively learn to solve more programming tasks.

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model
during the compilation phase:

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

\subsection{Exploration: Searching for Programs}\label{explorationSection}

@@

\subsection{Compilation: Learning a Recognition Model}\label{recognitionSection}

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}
\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\qquad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\subsection{Compression: Learning a Generative Model (a DSL)}\label{grammarInductionSection}

@@

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.

@@

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We perform an exploration cycle before each compression and compilation  cycle.
(2) On the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a generative model has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
%% (3) Because the frontiers can grow very large,
%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     %     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Compile}})
     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}

     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D},\theta,q$
\end{algorithmic}
\end{algorithm}
